Optimization of String Rewriting Operations for 3D Fractal Generation with Genetic Algorithms

### 2018-07-13 Optimization-of-String-Rewriting-Operations-for-3D-Fractal-Generation-with-Genetic-Algorithms/paper.tex
Small change of conclusions.
Some more fixes were done.

%----------------------------------------------------------------------------------------
%   Abstract
%----------------------------------------------------------------------------------------
\begin{abstract}
String rewriting is a modification of the idea for the context-free grammar. Modification consists in the fact that there is not separation on terminal and nonterminal symbols. Each symbol in string rewriting, is considered as nonterminal and it can produce longer sequence. By this way infinite structures are created as fractals. In a video presentation by Jack Hodkinson the use of 2D images instead of text symbol pixels is suggested. Each pixel (geometric square) is divided in nine sub-squares. The color of the pixel determine the pattern in which the nine squares are arranged. By this way each pixel gives a rule for subdivision of the containing area. Hodkinson gives a method, to setup the fractal, to reproduce a particular 2D image (for example, the glyph of the Greek letter $\pi$). This problem is well known in the fractals theory as Fractal Inverse Problem (FIP). It is an optimization problem, so a good option is the use of genetic algorithm (GA), to assemble set of rules, for string rewriting with 2D pixels. Series of experiments with the size of substitution matrix (not only 3x3, but also 2x2 and 4x4) are done by Hodkinson. Also are made series of experiments with different number of colors, starting with black/white and reaching the final successful experiment with thirty shades of blue. In this study the idea for string rewriting is extended in the 3D space and instead of pixels voxels are used with 3x3x3 substitution matrix. The study is only on virtual models but with the usage of industrial tomography, 3D objects can be scanned and with 3D color printer generated fractals can be printed.
\keywords{fractal inverse problem, genetic algorithms, string rewriting}
\end{abstract}

@@

All experiments are done in Java programming language with Apache Genetic Algorithms Framework \cite{apache01} as an optimization approach and Printing in 3D - JavaSCAD \cite{printing01} as 3D visualization tool. The source code used for all experiments is an open source and can be found at \cite{balabanov01} for the interested readers who would like to reproduce the experiments and results. Increasing the space dimensions from 2D to 3D greatly increases the complexity of the optimization process. Because of the increased complexity less ambitious shape than the Greek letter $\pi$ is taken, in our case it is a voxelized sphere (Fig. \ref{fig:pic02}). For visualization purposes the first color in the set of colors is accepted as transparent.

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.0\linewidth]{pic02}
  \caption{Voxelized sphere.}
\label{fig:pic02}
\end{figure}
\FloatBarrier

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.0\linewidth]{pic03}
  \caption{Steps taken in the optimization process.}
\label{fig:pic03}
\end{figure}
\FloatBarrier

As it is shown in Fig. \ref{fig:pic03} the optimization process converges much more slower in the 3D space than what was achieved by Hodkinson in the 2D space.

@@

The extension of the idea for fractal generation from 2D to 3D space shows that in 3D space the process is much more time consuming than in the 2D space. Some speed-ups are possible by parallel implementation of the genetic algorithm \cite{shonkwiler02}. Voxel based representation of 3D objects consumes a lot of computational resources and generated objects are with limited usefulness. A possible application is compression of 3D images, which is similar to the idea of 2D images compression \cite{vences01,albundi01}. Achieved results are promising for further research of the fractal nature of the world in which all we are living.

>>>>

%----------------------------------------------------------------------------------------
%   Abstract
%----------------------------------------------------------------------------------------
\begin{abstract}
String rewriting is a modification of the idea for the context-free grammar. Modification consists in the fact that there is not separation on terminal and nonterminal symbols. Each symbol in string rewriting, is considered as nonterminal and it can produce longer sequence. By this way infinite structures are created as fractals. In a video presentation by Jack Hodkinson the use of 2D images instead of text symbol pixels is suggested. Each pixel (geometric square) is divided in nine sub-squares. The color of the pixel determine the pattern in which the nine squares are arranged. By this way each pixel gives a rule for subdivision of the containing area. Hodkinson gives a method, to setup the fractal, to reproduce a particular 2D image (for example, the glyph of the Greek letter $\pi$). This problem is well known in the fractals theory as Fractal Inverse Problem (FIP). It is an optimization problem, so a good option is the use of genetic algorithm (GA), to assemble set of rules, for string rewriting with 2D pixels. Series of experiments with the size of substitution matrix (not only 3x3, but also 2x2 and 4x4) are done by Hodkinson. Also are made series of experiments with different number of colors, starting with black/white and reaching the final successful experiment with thirty shades of blue.

\keywords{fractal inverse problem, genetic algorithms, string rewriting}
\end{abstract}

@@

All experiments are done in Java programming language with Apache Genetic Algorithms Framework \cite{apache01} as an optimization approach and Printing in 3D - JavaSCAD \cite{printing01} as 3D visualization tool. The source code used for all experiments is an open source and can be found at \cite{balabanov01} for the interested readers who would like to reproduce the experiments and results. Increasing the space dimensions from 2D to 3D greatly increases the complexity of the optimization process. Because of the increased complexity less ambitious shape than the Greek letter $\pi$ is taken, in our case it is a voxelized sphere (Fig. \ref{fig:pic02}). The sphere is one of the simplest 3D objects and that is why it was selected as optimization target. For visualization purposes the first color in the set of colors (1 of 32 in this case) is accepted as transparent. The goal of the experiment is form a single side cube to find such 3x3x3 set of transitions which will lead to fractal generation of a target 3D object (in this case sphere).

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.8\linewidth]{pic02}
  \caption{Voxelized sphere as fractal optimization target object.}
\label{fig:pic02}
\end{figure}
\FloatBarrier

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.0\linewidth]{pic03}
  \caption{Intermediate snapshots during the optimization process.}
\label{fig:pic03}
\end{figure}
\FloatBarrier

The optimization process converges much more slower in the 3D space than what was achieved by Hodkinson in the 2D space. As it is shown in Fig. \ref{fig:pic03}, starting form a single cube it goes to the desired shape too slowly.

@@

In this study the idea for string rewriting is extended in the 3D space and instead of pixels voxels are used with 3x3x3 substitution matrix. The study is only on virtual models but with the usage of industrial tomography, 3D objects can be scanned and with 3D color printer generated fractals can be printed. The extension of the idea for fractal generation from 2D to 3D space shows that in 3D space the process is much more time consuming than in the 2D space. Some speed-ups are possible by parallel implementation of the genetic algorithm \cite{shonkwiler02}. Voxel based representation of 3D objects consumes a lot of computational resources and generated objects are with limited usefulness. A possible application is compression of 3D images, which is similar to the idea of 2D images compression \cite{vences01,albundi01}. Achieved results are promising for further research of the fractal nature of the world in which all we are living.
